The role of viscosity on drop impact forces

### 2024-08-31 Main/ViscousDropImpact_v10.tex
feat(drop-impact): Characterize the peak impact and take-off forces
The changes made in this commit focus on characterizing the peak impact force ($F_1$) and the peak take-off force ($F_2$) during the impact of a liquid drop on a non-wetting substrate. The key changes are:

1. Explained the two-peak structure of the normal force $F(t)$ observed in the authors' previous work, where the second peak coincides with a Worthington jet formation.
2. Conducted experiments, direct numerical simulations, and scaling analyses to understand the dependency of $F_1$ and $F_2$ on the Weber number ($We$) and

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{Figures/Schematic_v2.pdf}
	\caption{(a) Problem schematic with an axisymmetric computational domain used to study the impact of a drop with diameter $D_0$ and velocity $V_0$ on a non-wetting substrate. In the experiments, we use a quartz force sensor to measure the temporal variation of the impact force. The subscripts $d$ and $a$ denote the drop and air, respectively, to distinguish their material properties, which are the density $\rho$ and the dynamic viscosity $\eta$. The drop–air surface tension coefficient is $\gamma$. The grey dashed-dotted line represents the axis of symmetry, $r = 0$. Boundary air outflow is applied at the top and side boundaries (tangential stresses, normal velocity gradient, and ambient pressure are set to zero). The domain boundaries are far enough from the drop not to influence its impact process ($\mathcal{L}_\text{max} \gg D_0$, $\mathcal{L}_\text{max} = 8R$ in the worst case). (b) The phase space with control parameters: the Weber number ($We$: dimensionless kinetic energy) and the Ohnesorge number ($Oh$: dimensionless viscosity), exemplifying different applications.}
	\label{fig:schematic}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{Figures/Figure1_summary_v6.pdf}
	\caption{{\oo Comparison of the drop impact force $F(t)$ obtained from experiments and simulations for the three typical cases with impact velocity $V_0 = 1.2\,\si{\meter}/\si{\second}, 0.97\,\si{\meter}/\si{\second}, 0.96\,\si{\meter}/\si{\second}$, diameter $D_0 = 2.05\,\si{\milli\meter}, 2.52\,\si{\milli\meter}, 2.54\,\si{\milli\meter}$, surface tension $\gamma = 72\,\si{\milli\newton}/\si{\meter}, 61\,\si{\milli\newton}/\si{\meter}, 61\,\si{\milli\newton}/\si{\meter}$ and viscosity $\eta_d = 1\,\si{\milli\pascal\second}, 25.3\,\si{\milli\pascal\second}, 80.2\,\si{\milli\pascal\second}$. These parameter give $Oh = 0.0025, 0.06, 0.2$ and $We = 40$.
	For the three cases, the two peak amplitudes, $F_1/(\rho_dV_0^2D_0^2) \approx$ 0.82, 0.92, 0.99 at $t_1 \approx 0.03\sqrt{\rho_dD_0^3/\gamma}$ and $F_2/(\rho_dV_0^2D_0^2) \approx$ 0.37, 0.337, 0.1 at $t_2 \approx 0.42\sqrt{\rho_dD_0^3/\gamma}$, characterize the inertial shock from impact and the Worthington jet before takeoff, respectively. 
	The drop reaches the maximum spreading at $t_{\text{max}}$ when it momentarily stops and retracts until $0.8\sqrt{\rho_dD_0^3/\gamma}$ when the drop takes off ($F = 0$). The black and gray dashed lines in panel (a) mark $F = 0$ and the resolution $F = 0.5\,\si{\milli\newton}$ of our piezoelectric force transducer, respectively.
	(b) Four instances are further elaborated through numerical simulations for ($We = 40, Oh = 0.0025$), namely (i) $t = 0$ (touch-down), (ii) $t = 0.03\sqrt{\rho_dD_0^3/\gamma}$ ($t_1$), (iii) $t = 0.2\sqrt{\rho_dD_0^3/\gamma}$ ($t_{\text{max}}$), and (iv) $t = 0.42\sqrt{\rho_dD_0^3/\gamma}$ ($t_2$).
	The insets of panel (a) exemplify these four instances for the three representative cases illustrated here. The experimental snapshots are overlaid with the drop boundaries from simulations. 
	We stress the excellent agreement between experiments and simulations without any free parameters.
	The left part of each numerical snapshot shows (on a $\log_{10}$ scale) the dimensionless local viscous dissipation function $\tilde{\xi}_\eta \equiv \xi_\eta D_0/\left(\rho_dV_0^3\right) = 2Oh\left(\boldsymbol{\tilde{\mathcal{D}}:\tilde{\mathcal{D}}}\right)$, where $\boldsymbol{\mathcal{D}}$ is the symmetric part of the velocity gradient tensor, and the right part the velocity field magnitude normalized with the impact velocity. The black velocity vectors are plotted in the center of mass reference frame of the drop to clearly elucidate the internal flow.  Also see supplementary videos SM1-SM3.\bb}}
\label{fig:summary}
\end{figure}

@@

These forces have been studied by \citet{zhang2022impact}, employing experiments and simulations and deriving scaling laws. 
A liquid drop impacting a non-wetting substrate undergoes a series of phases--spreading, recoiling, and potentially rebounding \citep{chantelot2018rebonds}--driven by the normal reaction force exerted by the substrate (figure~\ref{fig:summary}). 
The moment of touch-down (see figures~\ref{fig:summary}a,b $t = 0$ to $t_1$) \citep{wagner1932stoss, Philippi2016, Gordillo2018} is not surprisingly associated with a pronounced peak in the temporal evolution of the drop impact force $F(t)$ owing to the sudden deceleration as high as $100$ times the acceleration due to gravity \citep{Clanet2004} (figure~\ref{fig:summary}a, \Rev{$F_1/(\rho_dV_0^2D_0^2) \approx$ 0.82, 0.92, 0.99 for $Oh = 0.0025, 0.06,\,\text{and}\,0.2$, respectively; at $t_1 \approx 0.03\sqrt{\rho_dD_0^3/\gamma}$}.
%$F(t = 0.37\,\si{\milli\second}) \approx 5.1\,\si{\milli\newton}$). 
The force diminishes as the drop reaches its maximum spreading diameter (figure~\ref{fig:summary}a,b $t = t_m$). 
\citet{zhang2022impact} revealed that also the jump-off is accompanied by a peak in the normal reaction force, which was up to then unknown (figure~\ref{fig:summary}a, 
%$F_2(t_2 = 4.63\,\si{\milli\second}) \approx 2.3\,\si{\milli\newton}$ 
\Rev{$F_2/(\rho_dV_0^2D_0^2) \approx$ 0.37, 0.337, 0.1 for $Oh = 0.0025, 0.06,\,\text{and}\,0.2$, respectively;
for the second force peak amplitude--at time $t_2 \approx 0.42\sqrt{\rho_dD_0^3/\gamma}$ after impact)}. The second peak in the force also coincides with the formation of a Worthington jet, a narrow upward jet of liquid that can form due to flow focusing by the retracting drop (figure~\ref{fig:summary}a,b $t = t_2$). Under certain conditions ($We \approx 9$, $Oh = 0.0025$), this peak can be even more pronounced than the first. This discovery is critical for superhydrophobicity which is volatile and can fail due to external disturbances such as pressure \citep{Lafuma2003, Callies2005, Sbragaglia2007, Li2017}, evaporation \citep{Tsai2010, Chen2012, Papadopoulos2013},  mechanical vibration \citep{Bormashenko2007}, or the impact forces of prior droplets \citep{Bartolo2006Bouncing}.

@@

In the experimental setup, shown schematically in figure~\ref{fig:schematic}(a), a liquid drop impacts a superhydrophobic substrate. For water drops, such a surface is coated with silanized silica nanobeads with a diameter of $20\,\si{\nano\meter}$ (Glaco Mirror Coat Zero; Soft99) resulting in the advancing and receding contact angles of $167 \pm 2^{\circ}$ and $154 \pm 2^{\circ}$, respectively \citep{Gauthier2015, Li2017}. On the other hand, for viscous aqueous glycerin drops, the upper surface is coated with an acetone solution of hydrophobic beads (Ultra ever Dry, Ultratech International, a typical bead size of 20 nm), resulting in the advancing and receding contact angles of $166 \pm 4^{\circ}$ and $159 \pm 2^{\circ}$, respectively \citep{Jha2020}. 
\Rev{The properties of the impacting drop are controlled using water-glycerol mixtures with viscosities $\eta_d$ varying by almost two orders of magnitude, from $1\,\si{\milli\pascal}\si{\second}$ to $80.2\,\si{\milli\pascal}\si{\second}$. Surface tension is either $72\,\si{\milli\newton}/\si{\meter}$ (pure water) or $61\,\si{\milli\newton}/\si{\meter}$ (glycerol), while density $\rho_d$ ranges from $1000\,\si{\kilo\gram}/\si{\cubic\meter}$ to $1220\,\si{\kilo\gram}/\si{\cubic\meter}$, as detailed in table~1 \citep{cheng2008formula, volk2018density, Jha2020}.
We note that using liquids such as silicone oil can provide a broader range of viscosity variation when paired with a superamphiphobic substrate \citep{deng2012candle}. Additionally, employing drops of smaller radii facilitates the exploration of higher Ohnesorge numbers ($Oh$, see~\eqref{Eq:OhDefinition}). 
The drop diameter $D_0$ is controlled between $2.05\,\si{\milli\meter}$ and $2.76\,\si{\milli\meter}$ by pushing it through a calibrated needle (see appendix~\ref{app:error} for details). 
Consequently, we calculate $Oh$ using the properties in table~\ref{tab:table00}.} 
The Weber number ($We$, see~\eqref{Eq:WeDefinition}) is set using the impact velocity $V_0$ varying between $0.38\,\si{\meter}/\si{\second}$ and $2.96\,\si{\meter}/\si{\second}$ by changing the release height of the drops above the substrate. 
All experiments are conducted at ambient pressure and temperature. The impact force is directly measured using a high-precision piezoelectric force transducer (Kistler 9215A) with a resolution of $0.5\,\si{\milli\newton}$. During these measurements, the high-frequency vibrations induced by the measurement system and the surrounding noise are spectrally removed using a low pass filter with a cut-off frequency of $5\,\si{\kilo\hertz}$, following the procedure in \citet{Li2014, Zhang2017, Gordillo2018, Mitchell2019}. 
The experiment also employs a high-speed camera (Photron Fastcam Nova S12) synchronized at $10,000$ fps with a shutter speed $1/20,000\,\si{\second}$. \Rev{Throughout the manuscript, the error bars are of statistical nature (one standard deviation) and originate from repeated trials. They are visible if they are larger than the marker size. We refer the readers to the supplementary material of \citet{zhang2022impact}  and appendix~\ref{app:error} for further details of the experimental setup and error characterization of the dimensionless control parameters.}

@@

We leverage the axial symmetry of the drop impact (figure~\ref{fig:schematic}a). This axial symmetry breaks at large $We$ ($\ge 100$ for water drops and even larger Weber number for more viscous drops), owing to destabilization by the surrounding gas after splashing \citep{xu2005drop, Eggers2010, Driscoll2011, riboux2014experiments, Josserand2016, zhang2022impact}. 
To solve the governing equations~\eqref{eqn:continuity}-\eqref{eqn:VoF}, the velocity field $\boldsymbol{v}$ and time $t$ are normalized by the inertio-capillary scales, $V_\gamma = \sqrt{\gamma/\rho_dD_0}$ and $\tau_\gamma = \sqrt{\rho_dD_0^3/\gamma}$, respectively. Furthermore, the pressure is normalized using the capillary pressure scale $p_\gamma = \gamma/D_0$. In such a conceptualization, $Oh$ and $We$ described in \S\ref{sec:intro} uniquely determine the system. 
The Ohnesorge number based on air viscosity $Oh_a = \left(\eta_a/\eta_d\right)Oh$ and air-drop density ratio $\rho_a/\rho_d$ are fixed at $10^{-5}$ and $10^{-3}$, respectively to minimize the influence of the surrounding medium on the impact forces. 
\Rev{Lastly, we keep the Bond number $Bo$ (see~\eqref{Eq:BoDefinition}) fixed at 1 throughout the manuscript. In our system, the relevance of gravity is characterized by the dimensionless Froude number $Fr = V_0^2/(gD_0) = We/Bo$ which compares inertia with gravity. Throughout this manuscript, $Fr > 1$ and gravity's role is sub-dominant compared to inertia (for detailed discussion, see appendix~\ref{app:gravity}).}
The substrate is modeled as a no-slip and non-penetrable wall, whereas vanishing stress and pressure are applied at the remaining boundaries to mimic outflow conditions for the surrounding air. The domain boundaries are far enough from the drop not to influence its impact process ($\mathcal{L}_\text{max} \gg D_0$, $\mathcal{L}_\text{max} = 8R$ in the worst case). 
At $t = 0$, in our simulations, we release a spherical drop whose south pole is $0.05D_0$ away from the substrate and is falling with a velocity $V_0$. 
\Rev{It is important to note that large experimental drops may deviate from perfect sphericity due to air drag as they fall after detaching from the needle and potential residual oscillations from detachment. These shape perturbations are more pronounced in cases with low Weber and Ohnesorge numbers. To quantify this non-sphericity, we measure the drop's aspect ratio (horizontal to vertical diameter) immediately before substrate contact. The precise pre-impact drop shape can significantly influence subsequent impact dynamics \citep{thoraval-2013-jfm, yun2017bouncing,Zhang2019}. In our experiments, we constrain our analysis to drops with aspect ratios between 0.96 and 1.05. Given this narrow range, we posit that the impact of these shape variations is negligible compared to the experimental error bars derived from repeated trials under identical nominal conditions.}
The simulations utilize adaptive mesh refinement to finely resolve the velocity, viscous dissipation, and the VoF tracer fields. A minimum grid size $\Delta = D_0/2048$ is used for this study.

@@

\noindent{\bf  Code availability\bf{.}} The codes used in the present article, \Rev{and the parameters and data to reproduce figures~\ref{fig:F1Anatomy} and~\ref{fig:F2Anatomy}} are permanently available at \citet{basiliskVatsal}.\\

@@

\oo
\section{Note on the error characterization for the control parameters}
\label{app:error}
\bb
\Rev{This appendix outlines the methodology for characterizing experimental errors in quantification of the drop's size and impact velocities which is crucial for accurate calculation of dimensionless control parameters, $We$ and $Oh$.
The drop diameter determination involves multiple steps. First, we measure the total mass ($M_{100}$) of 100 drops using an electric balance. From this mass, using the liquid density and assuming spherical shape, we calculated the drop diameter ($D_0$). 
We repeated this process five times, yielding $D_{0,1}$ through $D_{0,5}$. The average of these measurements provided the final drop diameter ($D_0$) and its standard error.
For impact velocity determination, we extracted data from experimental high-speed imagery. By tracking the drop center's position in successive frames prior to substrate contact, and knowing the frame rate, we calculated the impact velocity. We repeated this process for five trials, obtaining $V_{0,1}$ through $V_{0,5}$. The average of these values gave the final impact velocity ($V_0$) and its standard error.}

\oo
The standard errors for drop diameters do not exceed $0.13\,\si{\milli\meter}$. For instance, drops with Ohnesorge numbers of $0.0025$, $0.06$, and $0.2$ have diameters of $2.05 \pm 0.13\,\si{\milli\meter}$, $2.52 \pm 0.11\,\si{\milli\meter}$, and $2.54 \pm 0.09\,\si{\milli\meter}$, respectively. 
The standard errors for impact velocities did not exceed $0.02\,\si{\meter}/\si{\second}$. For the same $Oh$ values, the impact velocities were $1.2 \pm 0.002\,\si{\meter}/\si{\second}$, $0.97 \pm 0.01\,\si{\meter}/\si{\second}$, and $0.96 \pm 0.01\,\si{\meter}/\si{\second}$, respectively.
The combined errors in $D_0$ and $V_0$ resulted in approximately $\pm 7\%$ error in Weber number $We$ and $\pm 3\%$ error in Ohnesorge number $Oh$. Consequently, the horizontal error bars, which relate to errors in the control parameters, are smaller than the symbol sizes in our figures.
\bb

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{Figures/figureA1.pdf}
	\caption{{\oo Comparison of the drop impact force $F(t)$ obtained from simulations for the four different Bond numbers $Bo = 0, 0.5, 1, 1.25$. Here, $Oh = 0.06$ and $We = 40$. Both force peaks $F_1$ and $F_2$ as well as time to reach these peaks $t_1$ and $t_2$ are invariant to variation in $Bo$. \bb}}
	\label{fig:AppGravity}
\end{figure}

\oo 
\section{Role of gravity on drop impact forces}
\label{app:gravity}
\bb

\Rev{Following table~\ref{tab:table00} and considering the variation in impacting drop diameter (appendix~\ref{app:error}), the Bond number (equation~\eqref{Eq:BoDefinition}) in our experiments ranges from $0.5$ to $1.25$, introducing an additional dimensionless control parameter alongside $We$ and $Oh$. 
Gravity typically plays a negligible role in these impact processes \citep{sanjay_chantelot_lohse_2023,sanjay2024PRL}. We undertook a sensitivity test varying the Bond number from $0$ to $1.25$ in our simulations. 
Figure~\ref{fig:AppGravity} confirms the leading-order Bond invariance of the results as the impact force profiles, including both force peaks $F_1$ and $F_2$ and their corresponding times $t_1$ and $t_2$, remain invariant to these Bond number variations.
Notably, while gravity does play a role in drop impact dynamics, particularly for longer time scales and in determining the critical Ohnesorge number $Oh_c$ for bouncing inhibition (see figure~\ref{fig:RegimeMaps}b and \citet{sanjay_chantelot_lohse_2023}), its effect on the initial impact force peaks is minimal for the parameter range studied here (large Froude numbers, $Fr > 1$). This Bond number invariance allows us to focus on the more dominant effects of Weber and Ohnesorge numbers.
Consequently, we selected the representative value of $Bo = 1$, corresponding to a diameter of $0.00254\,\si{\milli\meter}$, density $1000\,\si{\kilogram/\meter^3}$, gravitational acceleration $10\,\si{\meter/\second^2}$, and surface tension $0.061\,\si{\newton/\meter}$.}

>>>>

	\begin{figure}
		\centering
		\includegraphics[width=\textwidth]{Figures/Schematic_v2.pdf}
		\caption{(a) Problem schematic with an axisymmetric computational domain used to study the impact of a drop with diameter $D_0$ and velocity $V_0$ on a non-wetting substrate. In the experiments, we use a quartz force sensor to measure the temporal variation of the impact force. The subscripts $d$ and $a$ denote the drop and air, respectively, to distinguish their material properties, which are the density $\rho$ and the dynamic viscosity $\eta$. The drop–air surface tension coefficient is $\gamma$. The grey dashed-dotted line represents the axis of symmetry, $r = 0$. Boundary air outflow is applied at the top and side boundaries (tangential stresses, normal velocity gradient, and ambient pressure are set to zero). The domain boundaries are far enough from the drop not to influence its impact process ($\mathcal{L}_\text{max} \gg D_0$, $\mathcal{L}_\text{max} = 8R$ in the worst case). (b) The phase space with control parameters: the Weber number ($We$: dimensionless kinetic energy) and the Ohnesorge number ($Oh$: dimensionless viscosity), exemplifying different applications.}
		\label{fig:schematic}
	\end{figure}
	\begin{figure}
		\centering
		\includegraphics[width=\textwidth]{Figures/Figure1_summary_v6.pdf}
		\caption{Comparison of the drop impact force $F(t)$ obtained from experiments and simulations for the three typical cases with impact velocity $V_0 = 1.2\,\si{\meter}/\si{\second}, 0.97\,\si{\meter}/\si{\second}, 0.96\,\si{\meter}/\si{\second}$, diameter $D_0 = 2.05\,\si{\milli\meter}, 2.52\,\si{\milli\meter}, 2.54\,\si{\milli\meter}$, surface tension $\gamma = 72\,\si{\milli\newton}/\si{\meter}, 61\,\si{\milli\newton}/\si{\meter}, 61\,\si{\milli\newton}/\si{\meter}$ and viscosity $\eta_d = 1\,\si{\milli\pascal\second}, 25.3\,\si{\milli\pascal\second}, 80.2\,\si{\milli\pascal\second}$. These parameter give $Oh = 0.0025, 0.06, 0.2$ and $We = 40$.
			For the three cases, the two peak amplitudes, $F_1/(\rho_dV_0^2D_0^2) \approx$ 0.82, 0.92, 0.99 at $t_1 \approx 0.03\sqrt{\rho_dD_0^3/\gamma}$ and $F_2/(\rho_dV_0^2D_0^2) \approx$ 0.37, 0.337, 0.1 at $t_2 \approx 0.42\sqrt{\rho_dD_0^3/\gamma}$, characterize the inertial shock from impact and the Worthington jet before takeoff, respectively. 
			The drop reaches the maximum spreading at $t_{\text{max}}$ when it momentarily stops and retracts until $0.8\sqrt{\rho_dD_0^3/\gamma}$ when the drop takes off ($F = 0$). The black and gray dashed lines in panel (a) mark $F = 0$ and the resolution $F = 0.5\,\si{\milli\newton}$ of our piezoelectric force transducer, respectively.
			(b) Four instances are further elaborated through numerical simulations for ($We = 40, Oh = 0.0025$), namely (i) $t = 0$ (touch-down), (ii) $t = 0.03\sqrt{\rho_dD_0^3/\gamma}$ ($t_1$), (iii) $t = 0.2\sqrt{\rho_dD_0^3/\gamma}$ ($t_{\text{max}}$), and (iv) $t = 0.42\sqrt{\rho_dD_0^3/\gamma}$ ($t_2$).
			The insets of panel (a) exemplify these four instances for the three representative cases illustrated here. The experimental snapshots are overlaid with the drop boundaries from simulations. 
			We stress the excellent agreement between experiments and simulations without any free parameters.
			The left part of each numerical snapshot shows (on a $\log_{10}$ scale) the dimensionless local viscous dissipation function $\tilde{\xi}_\eta \equiv \xi_\eta D_0/\left(\rho_dV_0^3\right) = 2Oh\left(\boldsymbol{\tilde{\mathcal{D}}:\tilde{\mathcal{D}}}\right)$, where $\boldsymbol{\mathcal{D}}$ is the symmetric part of the velocity gradient tensor, and the right part the velocity field magnitude normalized with the impact velocity. The black velocity vectors are plotted in the center of mass reference frame of the drop to clearly elucidate the internal flow.  Also see supplementary videos SM1-SM3.}
		\label{fig:summary}
	\end{figure}

@@

	These forces have been studied by \citet{zhang2022impact}, employing experiments and simulations and deriving scaling laws. 
	A liquid drop impacting a non-wetting substrate undergoes a series of phases--spreading, recoiling, and potentially rebounding \citep{chantelot2018rebonds}--driven by the normal reaction force exerted by the substrate (figure~\ref{fig:summary}). 
	The moment of touch-down (see figures~\ref{fig:summary}a,b $t = 0$ to $t_1$) \citep{wagner1932stoss, Philippi2016, Gordillo2018} is not surprisingly associated with a pronounced peak in the temporal evolution of the drop impact force $F(t)$ owing to the sudden deceleration as high as $100$ times the acceleration due to gravity \citep{Clanet2004} (figure~\ref{fig:summary}a, $F_1/(\rho_dV_0^2D_0^2) \approx$ 0.82, 0.92, 0.99 for $Oh = 0.0025, 0.06,\,\text{and}\,0.2$, respectively; at $t_1 \approx 0.03\sqrt{\rho_dD_0^3/\gamma}$.
	The force diminishes as the drop reaches its maximum spreading diameter (figure~\ref{fig:summary}a,b $t = t_m$). 
	\citet{zhang2022impact} revealed that also the jump-off is accompanied by a peak in the normal reaction force, which was up to then unknown (figure~\ref{fig:summary}a, 
	$F_2/(\rho_dV_0^2D_0^2) \approx$ 0.37, 0.337, 0.1 for $Oh = 0.0025, 0.06,\,\text{and}\,0.2$, respectively; for the second force peak amplitude--at time $t_2 \approx 0.42\sqrt{\rho_dD_0^3/\gamma}$ after impact)
	The second peak in the force also coincides with the formation of a Worthington jet, a narrow upward jet of liquid that can form due to flow focusing by the retracting drop (figure~\ref{fig:summary}a,b $t = t_2$). Under certain conditions ($We \approx 9$, $Oh = 0.0025$), this peak can be even more pronounced than the first. This discovery is critical for superhydrophobicity which is volatile and can fail due to external disturbances such as pressure \citep{Lafuma2003, Callies2005, Sbragaglia2007, Li2017}, evaporation \citep{Tsai2010, Chen2012, Papadopoulos2013},  mechanical vibration \citep{Bormashenko2007}, or the impact forces of prior droplets \citep{Bartolo2006Bouncing}.

@@

	In the experimental setup, shown schematically in figure~\ref{fig:schematic}(a), a liquid drop impacts a superhydrophobic substrate. For water drops, such a surface is coated with silanized silica nanobeads with a diameter of $20\,\si{\nano\meter}$ (Glaco Mirror Coat Zero; Soft99) resulting in the advancing and receding contact angles of $167 \pm 2^{\circ}$ and $154 \pm 2^{\circ}$, respectively \citep{Gauthier2015, Li2017}. On the other hand, for viscous aqueous glycerin drops, the upper surface is coated with an acetone solution of hydrophobic beads (Ultra ever Dry, Ultratech International, a typical bead size of 20 nm), resulting in the advancing and receding contact angles of $166 \pm 4^{\circ}$ and $159 \pm 2^{\circ}$, respectively \citep{Jha2020}. 
	The properties of the impacting drop are controlled using water-glycerol mixtures with viscosities $\eta_d$ varying by almost two orders of magnitude, from $1\,\si{\milli\pascal}\si{\second}$ to $80.2\,\si{\milli\pascal}\si{\second}$. Surface tension is either $72\,\si{\milli\newton}/\si{\meter}$ (pure water) or $61\,\si{\milli\newton}/\si{\meter}$ (glycerol), while density $\rho_d$ ranges from $1000\,\si{\kilo\gram}/\si{\cubic\meter}$ to $1220\,\si{\kilo\gram}/\si{\cubic\meter}$, as detailed in table~1 \citep{cheng2008formula, volk2018density, Jha2020}.
	We note that using liquids such as silicone oil can provide a broader range of viscosity variation when paired with a superamphiphobic substrate \citep{deng2012candle}. Additionally, employing drops of smaller radii facilitates the exploration of higher Ohnesorge numbers ($Oh$, see~\eqref{Eq:OhDefinition}). 
	The drop diameter $D_0$ is controlled between $2.05\,\si{\milli\meter}$ and $2.76\,\si{\milli\meter}$ by pushing it through a calibrated needle (see appendix~\ref{app:error} for details). 
	Consequently, we calculate $Oh$ using the properties in table~\ref{tab:table00}.
	The Weber number ($We$, see~\eqref{Eq:WeDefinition}) is set using the impact velocity $V_0$ varying between $0.38\,\si{\meter}/\si{\second}$ and $2.96\,\si{\meter}/\si{\second}$ by changing the release height of the drops above the substrate. 
	All experiments are conducted at ambient pressure and temperature. The impact force is directly measured using a high-precision piezoelectric force transducer (Kistler 9215A) with a resolution of $0.5\,\si{\milli\newton}$. During these measurements, the high-frequency vibrations induced by the measurement system and the surrounding noise are spectrally removed using a low pass filter with a cut-off frequency of $5\,\si{\kilo\hertz}$, following the procedure in \citet{Li2014, Zhang2017, Gordillo2018, Mitchell2019}. 
	The experiment also employs a high-speed camera (Photron Fastcam Nova S12) synchronized at $10,000$ fps with a shutter speed $1/20,000\,\si{\second}$. Throughout the manuscript, the error bars are of statistical nature (one standard deviation) and originate from repeated trials. They are visible if they are larger than the marker size. We refer the readers to the supplementary material of \citet{zhang2022impact}  and appendix~\ref{app:error} for further details of the experimental setup and error characterization of the dimensionless control parameters.

@@

	We leverage the axial symmetry of the drop impact (figure~\ref{fig:schematic}a). This axial symmetry breaks at large $We$ ($\ge 100$ for water drops and even larger Weber number for more viscous drops), owing to destabilization by the surrounding gas after splashing \citep{xu2005drop, Eggers2010, Driscoll2011, riboux2014experiments, Josserand2016, zhang2022impact}. 
	To solve the governing equations~\eqref{eqn:continuity}-\eqref{eqn:VoF}, the velocity field $\boldsymbol{v}$ and time $t$ are normalized by the inertio-capillary scales, $V_\gamma = \sqrt{\gamma/\rho_dD_0}$ and $\tau_\gamma = \sqrt{\rho_dD_0^3/\gamma}$, respectively. Furthermore, the pressure is normalized using the capillary pressure scale $p_\gamma = \gamma/D_0$. In such a conceptualization, $Oh$ and $We$ described in \S\ref{sec:intro} uniquely determine the system. 
	The Ohnesorge number based on air viscosity $Oh_a = \left(\eta_a/\eta_d\right)Oh$ and air-drop density ratio $\rho_a/\rho_d$ are fixed at $10^{-5}$ and $10^{-3}$, respectively to minimize the influence of the surrounding medium on the impact forces. 
	Lastly, we keep the Bond number $Bo$ (see~\eqref{Eq:BoDefinition}) fixed at 1 throughout the manuscript. In our system, the relevance of gravity is characterized by the dimensionless Froude number $Fr = V_0^2/(gD_0) = We/Bo$ which compares inertia with gravity. Throughout this manuscript, $Fr > 1$ and gravity's role is sub-dominant compared to inertia (for detailed discussion, see appendix~\ref{app:gravity}).
	The substrate is modeled as a no-slip and non-penetrable wall, whereas vanishing stress and pressure are applied at the remaining boundaries to mimic outflow conditions for the surrounding air. The domain boundaries are far enough from the drop not to influence its impact process ($\mathcal{L}_\text{max} \gg D_0$, $\mathcal{L}_\text{max} = 8R$ in the worst case). 
	At $t = 0$, in our simulations, we release a spherical drop whose south pole is $0.05D_0$ away from the substrate and is falling with a velocity $V_0$. 
	It is important to note that large experimental drops may deviate from perfect sphericity due to air drag as they fall after detaching from the needle and potential residual oscillations from detachment. These shape perturbations are more pronounced in cases with low Weber and Ohnesorge numbers. To quantify this non-sphericity, we measure the drop's aspect ratio (horizontal to vertical diameter) immediately before substrate contact. The precise pre-impact drop shape can significantly influence subsequent impact dynamics \citep{thoraval-2013-jfm, yun2017bouncing,Zhang2019}. In our experiments, we constrain our analysis to drops with aspect ratios between 0.96 and 1.05. Given this narrow range, we posit that the impact of these shape variations is negligible compared to the experimental error bars derived from repeated trials under identical nominal conditions.
	The simulations utilize adaptive mesh refinement to finely resolve the velocity, viscous dissipation, and the VoF tracer fields. A minimum grid size $\Delta = D_0/2048$ is used for this study.

@@

	\noindent{\bf  Code availability\bf{.}} The codes used in the present article, and the parameters and data to reproduce figures~\ref{fig:F1Anatomy} and~\ref{fig:F2Anatomy} are permanently available at \citet{basiliskVatsal}.\\

@@

	\section{Note on the error characterization for the control parameters}
	\label{app:error}

	This appendix outlines the methodology for characterizing experimental errors in quantification of the drop's size and impact velocities which is crucial for accurate calculation of the dimensionless control parameters $We$ and $Oh$.
	The drop diameter determination involves multiple steps. First, we measure the total mass ($M_{100}$) of 100 drops using an electric balance. From this mass, using the liquid density and assuming spherical shape, we calculated the drop diameter ($D_0$). 
	We repeated this process five times, yielding $D_{0,1}$ through $D_{0,5}$. The average of these measurements provided the final drop diameter ($D_0$) and its standard error.
	For impact velocity determination, we extracted data from experimental high-speed imagery. By tracking the drop center's position in successive frames prior to substrate contact, and knowing the frame rate, we calculated the impact velocity. We repeated this process for five trials, obtaining $V_{0,1}$ through $V_{0,5}$. The average of these values gave the final impact velocity ($V_0$) and its standard error.

	The standard errors for drop diameters do not exceed $0.13\,\si{\milli\meter}$. For instance, drops with Ohnesorge numbers of $0.0025$, $0.06$, and $0.2$ have diameters of $2.05 \pm 0.13\,\si{\milli\meter}$, $2.52 \pm 0.11\,\si{\milli\meter}$, and $2.54 \pm 0.09\,\si{\milli\meter}$, respectively. 
	The standard errors for impact velocities did not exceed $0.02\,\si{\meter}/\si{\second}$. For the same $Oh$ values, the impact velocities were $1.2 \pm 0.002\,\si{\meter}/\si{\second}$, $0.97 \pm 0.01\,\si{\meter}/\si{\second}$, and $0.96 \pm 0.01\,\si{\meter}/\si{\second}$, respectively.
	The combined errors in $D_0$ and $V_0$ resulted in approximately $\pm 7\%$ error in Weber number $We$ and $\pm 3\%$ error in Ohnesorge number $Oh$. Consequently, the horizontal error bars, which relate to errors in the control parameters, are smaller than the symbol sizes in our figures.

	\begin{figure}
		\centering
		\includegraphics[width=\textwidth]{Figures/figureA1.pdf}
		\caption{Comparison of the drop impact force $F(t)$ obtained from simulations for the four different Bond numbers $Bo = 0, 0.5, 1, 1.25$. Here, $Oh = 0.06$ and $We = 40$. Both force peaks $F_1$ and $F_2$ as well as time to reach these peaks $t_1$ and $t_2$ are invariant to variation in $Bo$.}
		\label{fig:AppGravity}
	\end{figure}

	\section{Role of gravity on drop impact forces}
	\label{app:gravity}

	Following table~\ref{tab:table00} and considering the variation in impacting drop diameter (appendix~\ref{app:error}), the Bond number (equation~\eqref{Eq:BoDefinition}) in our experiments ranges from $0.5$ to $1.25$, introducing an additional dimensionless control parameter alongside $We$ and $Oh$. 
	Gravity typically plays a negligible role in these impact processes \citep{sanjay_chantelot_lohse_2023,sanjay2024PRL}. We undertook a sensitivity test varying the Bond number from $0$ to $1.25$ in our simulations. 
	Figure~\ref{fig:AppGravity} confirms the leading-order Bond invariance of the results as the impact force profiles, including both force peaks $F_1$ and $F_2$ and their corresponding times $t_1$ and $t_2$, remain invariant to these Bond number variations.
	Notably, while gravity does play a role in drop impact dynamics, particularly for longer time scales and in determining the critical Ohnesorge number $Oh_c$ for bouncing inhibition (see figure~\ref{fig:RegimeMaps}b and \citet{sanjay_chantelot_lohse_2023}), its effect on the initial impact force peaks is minimal for the parameter range studied here (large Froude numbers, $Fr > 1$). This Bond number invariance allows us to focus on the more dominant effects of Weber and Ohnesorge numbers.
	Consequently, we selected the representative value of $Bo = 1$, corresponding to a diameter of $0.00254\,\si{\milli\meter}$, density $1000\,\si{\kilogram/\meter^3}$, gravitational acceleration $10\,\si{\meter/\second^2}$, and surface tension $0.061\,\si{\newton/\meter}$.
